### 2019-08-23 check_progress.Rmd
check progress via Rmd

---
title: "LAGOSUS depth progress"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(dplyr)
library(googledrive)
library(sf)
library(ggplot2)
library(jsta)
library(LAGOSNE)
library(assertr)
```

```{r data}
# drive_find(type = "spreadsheet")
if(!file.exists("data/backups/depth_log_all.csv")){
  drive_download("depth_log_all", type = "csv",
                 path = "data/backups/depth_log_all.csv", overwrite = TRUE)
}

dt_raw <- read.csv("data/backups/depth_log_all.csv", stringsAsFactors = FALSE)
```

```{r clean, warning=FALSE, message=FALSE}
# convert ft to m
dt <- dt_raw %>%
  mutate_at(vars(contains("depth")), as.numeric) %>%
  mutate(max_depth_m = case_when(
    is.na(max_depth_m) & !is.na(max_depth_ft) ~ max_depth_ft * 0.3048,
    TRUE ~ max_depth_m
  )) %>%
  mutate(mean_depth_m = case_when(
    is.na(mean_depth_m) & !is.na(mean_depth_ft) ~ mean_depth_ft * 0.3048,
    TRUE ~ mean_depth_m
  )) %>%
  dplyr::select(-mean_depth_ft, -max_depth_ft) %>%
  # assign GLNC to state(s)
  coordinatize("lat", "lon") %>%
  mutate(has_max = !is.na(max_depth_m) & nchar(max_depth_m) > 0) %>%
  st_join(jsta::usa_sf()) %>%
  mutate(state.x = case_when(
    state.x == "Western States GLNC" & !is.na(state.y) ~ state.y, 
    TRUE ~ state.x)) %>%
  dplyr::select(-state.y) %>% dplyr::rename(state = state.x) %>%
  dplyr::filter(state != "Western States GLNC")

# TODO: assertr the arg below
# which(dt$mean_depth_m == dt$max_depth_m)
```

```{r analysis}
# How many states have been touched?
# Of those states, how what is the response rate for max/mean depth?

frac <- dt %>%
  st_drop_geometry() %>%
  group_by(state) %>%
  # dplyr::filter(state == "ID") %>%
  summarize_at(vars(max_depth_m:mean_depth_m),
               function(x) round(sum(!is.na(x) & nchar(x > 0)) / length(x), 2)) %>%
  data.frame() %>% setNames(c("state", "max_frac", "mean_frac"))

n <- dt %>%
  st_drop_geometry() %>%
  group_by(state) %>%
  summarize_at(vars(max_depth_m:mean_depth_m),
               function(x) sum(!is.na(x))) %>%
  data.frame() %>% setNames(c("state", "max_n", "mean_n"))

res <- left_join(frac, n, by = "state") %>%
  arrange(desc(max_frac))
```

```{r tables}
knitr::kable(res)
knitr::kable(arrange(res, desc(max_n)))
```

```{r fun_facts, warning=FALSE, eval=TRUE}
# calculate total number of lakes we're working with here
lg <- lagosne_load()

paste0(sum(res$max_n), " lakes in not NE")

paste0((sum(res$max_n) + sum(!is.na(lg$lakes_limno$maxdepth))),
       " lakes in NE and not NE")

paste0(round(sum(res$max_n) / nrow(dt) * 100, 2),
       " percent of lakes with maxdepth not in NE")

paste0(round(sum(!is.na(lg$lakes_limno$maxdepth)) / nrow(lg$lakes_limno) * 100, 2),
       " percent of lakes with maxdepth in NE")
```

```{r map}
ggplot() +
  geom_sf(data = jsta::usa_sf()) +
  geom_sf(data = dt, aes(color = has_max), size = 0.6)

# mapview::mapview(dt_sf)
```