### 2019-12-29 figures/00_cutoffs.Rmd
add cutoffs figure

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 00_cutoffs, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
lg <- lagosus_load("locus")
dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
  dplyr::select(lagoslakeid, lake_maxdepth_m)

dt_plot <- lg$locus$locus_characteristics %>%
  left_join(dt) %>%
  mutate(area_class =
           smart_cut(lake_waterarea_ha, c(1, 4, 40, 80, 300,
                                          400, 1000, 20000, Inf),
             labels = ~paste(sep="-", thousand_k(.y[1]), thousand_k(.y[2])))
         ) %>% drop_na(area_class) %>% group_by(area_class) %>%
  mutate(prop_maxdepth = round(mean(!is.na(lake_maxdepth_m)), 2)) %>%
  add_tally() %>%
  distinct(area_class, prop_maxdepth, n) %>%
  arrange(area_class)

# dt_plot %>%
#   ggplot() +
#   geom_col(aes(y = n, x = area_class)) +
#   geom_text(aes(y = n, x = area_class, label = prop_maxdepth),
#             vjust = -0.5, size = 4) +
#   ylim(0, max(dt_plot$n) + 10000) + 
#   xlab("Area (ha)") +
#   ggtitle("Proportion max depth availability by \n lake area class")

dt_plot %>%
  ggplot() +
  geom_point(aes(y = prop_maxdepth, x = area_class)) +
  # ylim(0, max(dt_plot$n) + 10000) + 
  ylim(1, 0) +
  xlab("Area (ha)") +
  ylab("Max depth availability (prop)") + 
  theme(axis.text.x = element_text(angle = 90))
```