### 2019-12-29 figures/01_heatmap.Rmd
move heatmap to separate file

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 01_heatmap, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
library(corrr)
library(pheatmap)

dt_raw_ne <- read.csv("../data/lagosne_depth_predictors.csv", 
                      stringsAsFactors = FALSE)

pretty_names <- gsub("mbgconhull_", "", names(dt_raw_ne))
pretty_names <- gsub("lake_", "", pretty_names)
pretty_names <- gsub("_nounits", "", pretty_names)
pretty_names <- gsub("focallake", "", pretty_names)

test <- dt_raw_ne %>%
  setNames(pretty_names) %>%
  dplyr::select_if(is.numeric) %>%
  dplyr::select(-lagoslakeid, -contains("decdeg"),
                -contains("nhd"),
                -contains("lagoslakeid"),
                -huc12) %>%
  correlate() %>%
  data.frame()
test_rowname    <- test$rowname
test            <- dplyr::select(test, -rowname)
row.names(test) <- test_rowname

pheatmap(t(test),
        color = colorRampPalette(
          RColorBrewer::brewer.pal(n = 7, name = "Reds"))(100))
```