### 2020-01-07 figures/01_hypsography.Rmd
improved title

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 01_hypsography, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}

ids <- factor(c("convex", "concave"))
polys <- data.frame(
  id = rep(ids, each = 3),
  x = c(0, 100, 0, 
        0, 100, 100), 
  y = c(100, 0, 0, 
        100, 100, 0))

dt_raw <- read.csv("../data/00_hypso/hypso.csv", stringsAsFactors = FALSE)

ggplot(data = dt_raw) +
  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
  geom_abline(intercept = -100, slope = 1, color = "red") +
  ylim(c(100, 0)) +
  ylab("Depth (percent)") + xlab("Area (percent)") + 
  ggtitle(paste0("Normalized hypsography for ", length(unique(dt_raw$llid)), " lakes")) + 
  theme(legend.position = c(0.84, 0.2), 
        legend.title = element_blank()) +
  facet_wrap(~state)

```