### 2020-01-08 figures/02_hypso_model.Rmd
add figures showing initial modelling results

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 02_hypso_model, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}

dt <- readRDS("../data/01_hypso_model/hypso_model.rds")
```

```{r 02_hypso_model_fitted, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=4}

# names(dt$ranger)
# class(dt$ranger$confusion)

dt_sub <- data.frame(dt$ranger$confusion) %>%
  mutate(goodbad = ifelse(data.frame(dt$ranger$confusion)$response == 
                      data.frame(dt$ranger$confusion)$truth, "good", "bad")) %>%
  group_by(truth) %>%
  mutate(prop = Freq/sum(Freq))

gg_ranger <- ggplot(data = dt_sub, 
                    aes(x = truth, y = response, 
                        fill = goodbad, alpha = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_manual(values = c(good = "green", bad = "red")) +
  theme_bw() + 
  ggtitle(paste0("Random Forest (acc=", 
                 round(dt$ranger$score, 2), ")"))

dt_sub <- data.frame(dt$rpart$confusion) %>%
  mutate(goodbad = ifelse(data.frame(dt$rpart$confusion)$response == 
                      data.frame(dt$rpart$confusion)$truth, "good", "bad")) %>%
  group_by(truth) %>%
  mutate(prop = Freq/sum(Freq))

gg_rpart <- ggplot(data = dt_sub, 
                    aes(x = truth, y = response, 
                        fill = goodbad, alpha = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_manual(values = c(good = "green", bad = "red")) +
  theme_bw() + 
  ggtitle(paste0("Recursive Partitioning (acc=", 
                 round(dt$rpart$score, 2), ")"))

theme_opts <- theme(legend.position = "none", 
                    plot.title = element_text(size = 8.5))
cowplot::plot_grid(gg_rpart + theme_opts, 
                   gg_ranger + theme_opts,
                   rel_heights = c(1, 0.1)
                   )

```

```{r 02_hypso_model_importance, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=4}

# names(dt$rpart)
data.frame(importance = dt$rpart$importance, stringsAsFactors = FALSE) %>% 
  mutate(variable = row.names(.)) %>%
  mutate(variable = forcats::fct_reorder(variable, importance)) %>%
  arrange(desc(importance)) %>%
  ggplot() + 
  geom_col(aes(x = variable, y = importance)) +
  coord_flip() +
  ggtitle("Recursive Partitioning Importance")

```