### 2020-02-06 manuscript/data.Rmd
split data viz and ms viz, closes #22

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{colortbl}
---

\arrayrulecolor{white}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6, dev = c("pdf"))
```

```{r maps_dot, fig.cap = "Map of lake maximum depth in the continental United States.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/00_map-1.pdf")
```

```{r maps_state_all, fig.cap = "Proportion of ALL lakes greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/00_map_state-1.pdf")
```

```{r maps_state_limno, fig.cap = "Proportion of lakes with limno data greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/00_map_state-2.pdf")
```

```{r cutoffs, fig.cap = "Proportion max depth availability by lake area class.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/00_cutoffs-1.pdf")
```

```{r contrasts_depth, fig.cap = "Lake maximum depth by categorical variables.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/01_contrasts_depth-1.pdf")
```

```{r contrasts_tally, fig.cap = "Number of lake depth measurements by categories.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/01_contrasts_tally-1.pdf")
```

```{r 01_programs_tally, fig.cap = "Number of programs by categories.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/01_programs_tally-1.pdf")
```