### 2020-02-06 figures/00_qa.Rmd
add figure showing variation among multiple data sources, closes #17

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 4, fig.height = 4)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 00_qa, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=2.5}
# variation among measurements for the same lake

dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
  group_by(lagoslakeid) %>%
  add_tally() %>%
  dplyr::filter(n > 1) %>%
  # dplyr::filter(lagoslakeid == 5426)
  summarize(# lake_maxdepth_std = sd(lake_maxdepth_m), 
         lake_maxdepth_range = range(lake_maxdepth_m)[2] - 
           range(lake_maxdepth_m)[1], 
         lake_maxdepth_prop = lake_maxdepth_range / range(lake_maxdepth_m)[2]) %>%
  setNames(c("lagoslakeid", 
             "Range", "Range / largest"))
# View(dplyr::arrange(dt, desc(Range)))
# plot(dt$Range, dt$`Range / largest`)

ggplot(data = tidyr::pivot_longer(dt, -lagoslakeid)) +
  geom_histogram(aes(x = value)) +
  facet_wrap(~name, scales = "free") +
  xlab("meters")
```

```{r lgne-area_vs_lgus-area, eval=FALSE}
lg    <- lagosus_load("locus")
lg_ne <- lagosne_load()
dt    <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE)

names(lg$locus$locus_link$l)
res <- dt %>% 
  left_join(
    dplyr::select(lg$locus$locus_link, lagoslakeid, lagosne_lagoslakeid)) %>%
  left_join(
    dplyr::select(lg$locus$locus_characteristics, 
                  lagoslakeid, lake_totalarea_ha)) %>%
  dplyr::filter(!is.na(lagosne_lagoslakeid)) %>%
  dplyr::filter(lagoslakeid == lagosne_lagoslakeid) %>%
  dplyr::filter(!is.na(lake_maxdepth_m)) %>%
  distinct(lagoslakeid, .keep_all = TRUE) %>%
  dplyr::select(lagoslakeid, lagosne_lagoslakeid, lake_waterarea_ha, 
                lake_totalarea_ha, lake_maxdepth_m) %>%
  left_join(dplyr::select(lg_ne$locus, lagosne_lagoslakeid = lagoslakeid, 
                          lake_area_ha)) %>%
  mutate(diff = lake_area_ha - lake_waterarea_ha) %>%
  arrange(desc(diff))

res2 <- res %>%
  dplyr::filter(abs(diff) < 0.1)
hist(res2$diff)

res3 <- res %>%
  dplyr::filter(abs(diff) > 1)
i <- 2
mapview(LAGOSUSgis::query_lake_poly(1222))
dplyr::filter(lg$locus$locus_characteristics, lagoslakeid == 14794)
dplyr::filter(dt, lagoslakeid == 14794)
dplyr::filter(lg_ne$locus, lagoslakeid == 14794)

+
  mapview(LAGOSNEgis::query_gis(
    "LAGOS_NE_All_Lakes_4ha", "lagoslakeid", res3$lagoslakeid[i]))

mapview(LAGOSNEgis::query_gis(
    "LAGOS_NE_All_Lakes_4ha", "lagoslakeid", 61727))

```

```{r misc, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide'}
threshold <- 5

dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
  dplyr::filter(!is.na(lake_maxdepth_m)) %>%
  group_by(lagoslakeid) %>%
  add_tally() %>%
  dplyr::filter(n >= 2) %>%
  mutate(diff = (max(lake_maxdepth_m) - min(lake_maxdepth_m)) / 
           max(lake_maxdepth_m)) %>%
  arrange(desc(diff))

%>%
  dplyr::filter(!(lagos_effort == "NLA" & diff > threshold)) %>%
  dplyr::filter(!(lagos_effort == "LAGOSNE" & diff > threshold))

dt %>% View()
distinct(dt, lagoslakeid, diff) %>%
  pull(diff) %>% hist(n = 100)

mapview(LAGOSUSgis::query_lake_poly(365518))

r <- raster("../data/nh_bathy/5787.tif")

ggplot() + geom_blank() + 
  theme(panel.background = element_blank())
LAGOSNE::lake_info(c(5592, 5787)) %>%
  dplyr::select(lagoslakeid, lagosname1, maxdepth, gnis_name, 
                lake_area_ha) %>%
  t() %>%
  gridExtra::grid.table()

hist(dt$diff, xlim = c(0, 10), n = 100)
median(dt$diff, na.rm = TRUE)

res_all <- read.csv("../data/00_bathy_depth/bathy_geometry.csv",
                    stringsAsFactors = FALSE)
dt_raw_ne <- read.csv("../data/lagosne_depth_predictors.csv",
                      stringsAsFactors = FALSE) %>%
  mutate(llid = as.character(lagoslakeid))
test      <- left_join(res_all, dt_raw_ne) %>%
  data.frame()

test$calc_depth_lgne <- calc_depth(test$buffer100m_slope_max, test$dist_deepest,
                                  grain = 10)
test$calc_depth_lgus <- calc_depth(test$inlake_slope, test$dist_deepest)
plot(test$maxdepth, test$calc_depth_lgus)
# plot(test$lake_maxdepth_m, test$calc_depth_lgus)
# abline(0, 1)
#
# ggplot(data = test) +
#   geom_point(aes(x = lake_maxdepth_m, y = calc_depth_lgus)) +
#   geom_abline(aes(slope = 1, intercept = 0)) +
#   facet_wrap(~state)

# plotly::ggplotly())

#
# hist(test$inlake_slope)
# hist(test$dist_deepest)
# hist(test$calc_depth_lgus)
#
# plot(test$calc_depth_lgne, test$calc_depth_lgus)
# plot(test$lake_maxdepth_m, test$calc_depth_lgne)

#
# # 2654
#
#
#
# arrange(res, desc(dist_between)) %>%
#   View()
#
# hist(res$dist_between)
# plot(res$dist_deepest, res$dist_viscenter)
# abline(0, 1)

# pnt_surface  <- st_point_on_surface(dt_poly)
## point furthest from shore
# test <- as_Spatial(dt_poly)
# r <- rasterize(SpatialPolygons(test@polygons), dt)
# r[is.na(r)] <- 0
# r[r == 1] <- NA
# test2 <- distance(r)
# which.max(test2)
# xy           <- xyFromCell(dt, which.max(test2[]))
# pnt_furthest  <- st_sfc(st_multipoint(xy), crs = st_crs(dt))
# pnt_centroid <- st_centroid(dt_poly)

# dt           <- raster("data/mn_bathy/2419.tif")
# compare_deepest_center(dt)
# mapview::mapview(dt_poly) +
#   mapview::mapview(pnt_viscenter) +
# mapview::mapview(r) +
#   mapview::mapview(pnt_deepest)

```