### 2020-02-11 manuscript/data.Rmd
test fig size caption, closes #25

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{colortbl}
---

\arrayrulecolor{white}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6, dev = c("pdf"))
knitr::opts_knit$set(eval.after = "fig.cap")
```

\pagebreak

![Summary of lake depth measurements](../tables/00_data.pdf){width=65%}

```{r maps_dot, fig.cap = paste0("Map of lake maximum depth in the continental United States (", paste0(system("fsizemb ../figures/00_map-1.pdf", intern = TRUE), " MB)")), warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/00_map-1.pdf")
# 
```

```{r maps_state_all, fig.cap = "Proportion of ALL lakes greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/00_map_state-1.pdf")
```

```{r maps_state_limno, fig.cap = "Proportion of lakes with limno data greater than 4ha with maximum depth availability by state.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/00_map_state-2.pdf")
```

```{r cutoffs, fig.cap = "Proportion max depth availability by lake area class.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/00_cutoffs-1.pdf")
```

```{r contrasts_depth, fig.cap = "Lake maximum depth by categorical variables.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/01_contrasts_depth-1.pdf")
```

```{r contrasts_tally, fig.cap = "Number of lake depth measurements by categories.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/01_contrasts_tally-1.pdf")
```

```{r 01_programs_tally, fig.cap = "Number of programs by categories.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/01_programs_tally-1.pdf")
```

```{r 00_qa, fig.cap = "Range of depth measurements for lakes with data from multiple sources.", warning=FALSE, echo = FALSE, eval=TRUE}
knitr::include_graphics("../figures/00_qa-1.pdf")
```