### 2020-04-08 manuscript/appendix.Rmd
format appendix for WRR

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{colortbl}
- \usepackage{float}
- \usepackage{caption}
---

\def\figurename{Figure S\hspace{-0.3em}}
\def\tablename{Table S\hspace{-0.3em}}

\arrayrulecolor{white}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", dev = c("pdf"))
knitr::opts_chunk$set(fig.pos = 'H')
```

\pagebreak

\centering
\includegraphics{../../../../JournalSubmissions/logos/agu_publications.png}

\emph{Water Resources Research}

\vspace{1em}

Supporting information for

\vspace{1em}

\textbf{Bathymetry data from thousands of lakes show that lake depth prediction is confounded by difficulty modeling inlake slope}

\vspace{1em}

J.Stachelek\textsuperscript{1},

\vspace{1em}

\textsuperscript{1}Department of Fisheries and Wildlife, Michigan State University, 480 Wilson Rd., East Lansing, Michigan 48824 USA

\vspace{8em}

\textbf{Contents of this file}

\vspace{4em}

\textbf{Introduction}

\pagebreak
\setcounter{figure}{0}

```{r maps_dot_bathy, fig.cap = "Map of study lakes showing A) lake maximum depth measurements, B) cross-section shape class, and C) reservoir classification.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
knitr::include_graphics("../figures/00_map_bathy-1.pdf")
```

```{r contrasts, fig.cap = "Lake characteristics by categorical variables.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
knitr::include_graphics("../figures/01_contrasts-1.pdf")
```

```{r hypsography, fig.cap = "Hypsography classification by state.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
knitr::include_graphics("../figures/01_hypsography-1.pdf")
```

```{r metric_correlations, fig.cap = "Pearson correlation coefficients of map-derived metrics against inlake slope and deepest point distance metrics.", warning=FALSE, echo = FALSE, eval=FALSE, out.width="11cm", fig.align='center', out.extra=''}
knitr::include_graphics("../figures/metric_correlations-1.pdf")
```

```{r heatmap, fig.cap = "Correlation matrix heatmap.", warning=FALSE, echo = FALSE, eval=FALSE}
knitr::include_graphics("../figures/slope-compare-1.pdf")
```

```{r glaciated_vs_shapeclass, fig.cap = "Cross tabulation between lake glaciation and shape class. Circles are proportional to the number of lakes in each category. If a lake has been glaciated it is more likely to be convex. However, if a lake has not been glaciated there is no tendencey towards a particular shape class.", warning=FALSE, echo = FALSE, eval=FALSE}
knitr::include_graphics("../figures/glaciated_vs_shapeclass-1.pdf")
```

```{r contrasts_tally, fig.cap = "Number of lake depth measurements by categories.", warning=FALSE, echo = FALSE, eval=FALSE}
knitr::include_graphics("../figures/01_contrasts_tally-1.pdf")
```

```{r 02_depth_model_fitted, fig.cap = "A model that does not include shape class over predicts depth in concave lakes but does not systematically under predict depth in convex lakes.", warning=FALSE, echo = FALSE, eval=FALSE}
knitr::include_graphics("../figures/02_depth_model_fitted-1.pdf")
```

```{r 02_depth_oliver-2, fig.cap = "Reproduction of Oliver 2016 using LAGOSUS data in LAGOSNE footprint.", warning=FALSE, echo = FALSE, eval=FALSE}
knitr::include_graphics("../figures/02_depth_oliver-2.pdf")
```

```{r 02_depth_oliver-1, fig.cap = "Reproduction of Oliver 2016 using LAGOSUS data in LAGOSNE footprint.", warning=FALSE, echo = FALSE, eval=FALSE}
knitr::include_graphics("../figures/02_depth_oliver-1.pdf")
```

```{r 02_depth_oliver-4, fig.cap = "Oliver 2016 observed versus predicted.", warning=FALSE, echo = FALSE, eval=FALSE}
knitr::include_graphics("../figures/02_depth_oliver-4.pdf")
```

```{r 02_hypso_model_fitted, fig.cap = "Confusion matrix comparing two classification methods", warning=FALSE, echo = FALSE, eval=FALSE}
knitr::include_graphics("../figures/02_hypso_model_fitted-1.pdf")
```

```{r 02_hypso_model_importance, fig.cap = "Recusive Partitioning variable importance.", warning=FALSE, echo = FALSE, eval=FALSE, out.extra=''}
knitr::include_graphics("../figures/02_hypso_model_importance-1.pdf")
```

<!-- ![Messager et al. 2016 equation table](../figures/messager_eq_table.png){width=100%} -->

```{r 01_geometry_map, fig.cap = "Green points show the 'visual center' of the lake. Red points show the true deepest point of a lake.", warning=FALSE, echo = FALSE, eval=FALSE}
knitr::include_graphics("../figures/gg_distance-3.pdf")
```