### 2020-11-09 figures/01_nearshore.Rmd
stash nearshore exploration figures

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 01_nearshore, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide', fig.height=3.5}
# set llid
flist <- list.files("../data/elevatr", pattern = "\\d*.tif",
                    full.names = TRUE, include.dirs = TRUE)
ll_id           <- 2294
max_buffer_dist <- 100 # to match Hollister (2011)

bathy_pnts <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
ll_poly    <- query_gis("LAGOS_NE_All_Lakes_4ha", "lagoslakeid", ll_id)
ll_iws     <- query_gis("IWS", "lagoslakeid", ll_id)
elev       <- raster(flist[grep(ll_id, flist)])

deepest_pnt         <- dplyr::filter(bathy_pnts, llid == ll_id)
r                   <- raster(paste0("../data/", 
                                     tolower(deepest_pnt$state), "_bathy/", ll_id, ".tif"))
st_crs(deepest_pnt) <- st_crs(r)

elev    <- projectRaster(elev, crs = st_crs(r)$proj4string)
ll_poly <- st_transform(ll_poly, st_crs(r))
ll_iws  <- st_transform(ll_iws, st_crs(r))

# approximates functions in lakemorpho package
elev      <- mask(elev, as_Spatial(ll_iws))
elev      <- mask(elev, as_Spatial(st_buffer(ll_poly, max_buffer_dist)))
elev <- crop(elev, as_Spatial(st_as_sfc(st_bbox(st_buffer(ll_poly, 100)))))
slope     <- terrain(elev, "slope")

slope_mean <- mean(slope@data@values, na.rm = TRUE) * res(elev)[1]
slope_max  <- max(slope@data@values, na.rm = TRUE) * res(elev)[1]

# pull slope on-line from deepest point
shore_pnt <- sf::st_nearest_points(deepest_pnt, st_transform(
  st_cast(ll_poly, "MULTILINESTRING"), st_crs(r)))
ll_buf <- st_buffer(ll_poly, 100)

buffer_line <- sf::st_nearest_points(shore_pnt,
                               st_transform(st_cast(ll_buf, "MULTILINESTRING"), st_crs(r)))
slope_online <- raster::extract(slope, as_Spatial(buffer_line))
slope_online <- mean(slope_online[[1]], na.rm = TRUE) * res(slope)[2]

res <- list(slope_mean = slope_mean, slope_max = slope_max,
     slope_online = slope_online) %>%
  bind_rows()

ggplot() + 
  geom_sf(data = st_transform(ll_buf, st_crs(r)), size = 1.5) +
  geom_sf(data = ll_poly, size = 1.5) +
  geom_raster(data = as.data.frame(terrain(r, "slope"), xy = TRUE),
              aes(x = x, y = y, fill = slope)) +
  geom_raster(data = as.data.frame(slope, xy = TRUE),
              aes(x = x, y = y, fill = slope)) +
  scale_fill_gradient(na.value = NA) +
  geom_sf(data = deepest_pnt) +
  geom_sf(data = shore_pnt) +
  geom_sf(data = buffer_line, color = "red") +
  theme_void()
```

```{r slope_compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5}

# load csv file from 00_get_nearshore
dt <- read.csv("data/00_geometry/nearshore.csv", 
               stringsAsFactors = FALSE) %>%
    dplyr::filter(inlake_slope < 0.25) %>%
  dplyr::filter(slope_online_mean < 0.6)

cowplot::plot_grid(
  ggplot(data = dt) + 
    geom_point(aes(x = slope_mean, y = slope_online_mean)) +
    geom_abline(aes(slope = 1, intercept = 0)), 
   ggplot(data = dt) + 
    geom_point(aes(x = slope_mean, y = inlake_slope)) +
    geom_abline(aes(slope = 1, intercept = 0)), 
   ggplot(data = dt) + 
    geom_point(aes(x = slope_online_mean, y = inlake_slope)) +
    geom_abline(aes(slope = 1, intercept = 0))
)

plot(dt$inlake_slope, dt$slope_mean,
     xlim = c(0, 2), ylim = c(0, 2))
plot(dt$inlake_slope, dt$slope_online,
     xlim = c(0, 2), ylim = c(0, 2))

# plot(dt$slope_mean, dt$slope_online)
abline(0, 1)

cor.test(dt$slope_mean, dt$inlake_slope)
cor.test(dt$slope_online, dt$inlake_slope)

hist(dt$inlake_slope / dt$slope_online)

```