### 2020-11-10 figures/01_geometry.Rmd
prototype Figure 2 best-fit lines, #20

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 01_geometry, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5}
# example map of true deepest point versus visual center

# setwd("figures")
dt         <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
hypso_pred <- read.csv("../data/00_hypso/hypso_predictors.csv", 
                       stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class, shape_class), 
                                  by = c("llid" = "lagoslakeid"))

# plot of true deepest versus visual centers
gg_scatter_state <- ggplot(data = st_drop_geometry(dt)) + 
  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = state)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  xlab("Deepest point distance (m)") +
  ylab("Visual center distance (m)")

gg_scatter_shape <- st_drop_geometry(dt) %>%
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::filter(dist_viscenter >= dist_deepest) %>%
  dplyr::filter(dist_deepest < 3200) %>%
  dplyr::arrange(desc(factor(shape_class, 
                        levels = c("neither", "concave", "convex")))) %>%
  ggplot(aes(x = dist_deepest, y = dist_viscenter, color = shape_class)) + 
  geom_point(size = 1) +
  # geom_abline(aes(intercept = 0, slope = 1)) +
  labs(color = "") +
  xlab("Deepest point distance (m)") +
  ylab("Distance to \n lake center (m)")

gg_scatter_reservoir <- st_drop_geometry(dt) %>%
  dplyr::filter(!is.na(reservoir_class)) %>%
  dplyr::filter(dist_viscenter >= dist_deepest) %>%
  ggplot() + 
  geom_point(aes(x = dist_deepest, y = dist_viscenter, color = reservoir_class), 
             size = 1) +
  scale_color_discrete(labels = c("NL     ", "Res")) + 
  # geom_abline(aes(intercept = 0, slope = 1)) +
  xlab("Deepest point distance (m)") +
  ylab("Distance to \n lake center (m)")
```

```{r gg_distance, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3}
gg_distance <- cowplot::plot_grid(gg_scatter_shape + 
                     theme(axis.title.x = element_text(color = "white")), 
                   gg_scatter_reservoir + 
                     labs(color = "") +
                     theme(axis.title.y = element_blank(), 
                           axis.text.y = element_blank()), 
                   rel_widths = c(1, 0.8))
gg_distance

gg_pntrng_deepest <- st_drop_geometry(dt) %>%
    dplyr::filter(!is.na(reservoir_class)) %>%
  group_by(reservoir_class) %>%
  summarize(dist_deepest = list(tibble::enframe(
    quantile(dist_deepest, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = reservoir_class, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Deepest point distance (m)") + xlab("")

gg_pntrng_viscenter <- st_drop_geometry(dt) %>%
    dplyr::filter(!is.na(reservoir_class)) %>%
  group_by(reservoir_class) %>%
  summarize(dist_viscenter = list(tibble::enframe(
    quantile(dist_viscenter, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = reservoir_class, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Distance to lake center (m)") + xlab("")

cowplot::plot_grid(gg_pntrng_deepest, gg_pntrng_viscenter)

get_pnt_map <- function(x){
  # x <- 3
  # x <- which(dt$llid == 257862)
  lg_example <- LAGOSUSgis::query_gis(
    "LAGOS_US_All_Lakes_1ha", "lagoslakeid", dt$llid[x])
  crs_example <- st_crs(
    raster(list.files(paste0("../data/", tolower(dt$state[x]), "_bathy"), 
               include.dirs = TRUE, full.names = TRUE)[1])
    )
  st_crs(dt$pnt_deepest)   <- crs_example
  st_crs(dt$pnt_viscenter) <- crs_example

  ggplot() + 
    geom_sf(data = lg_example) +
    geom_sf(data = dt$pnt_deepest[x], color = "red") + 
    geom_sf(data = dt$pnt_viscenter[x], color = "green") +
    theme_void()
}

# get_pnt_map(which(dt$llid == 5636))
# get_pnt_map(which(dt$llid == 257862))
# cowplot::plot_grid(get_pnt_map(which(dt$llid == 3791)) +
#                      ggtitle(3791), 
#                    get_pnt_map(which(dt$llid == 2772)) +
#                      ggtitle(2772), 
#                    get_pnt_map(which(dt$llid == 2655)) +
#                      ggtitle(2655), 
#                    get_pnt_map(which(dt$llid == 1822)) +
#                      ggtitle(1822)
#                    )

```

```{r slope-compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5}
# compare inlake slope, buffer slope, and online slope
# setwd("figures") 
bp <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
  mutate(llid = as.character(llid))
dt <- left_join(dt, st_drop_geometry(bp), 
                by = "llid") %>%
  dplyr::filter(!is.na(inlake_slope_mean))

# jsta::jheatmap
jheatmap(
  dplyr::select(dt, 
                slope_online_mean,
                inlake_slope_mean,
                slope_mean, 
                slope_mean_ne,
                dist_deepest, 
                dist_viscenter, 
                maxdepth,
                meandepth))

GGally::ggpairs(dplyr::select(dt, 
                inlake_slope_mean, inlake_slope.y,
                slope_online_mean,
                slope_mean, 
                # dist_deepest, 
                # dist_viscenter, 
                maxdepth), 
                progress = FALSE, 
                ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1), 
          strip.text.y = element_text(angle = 0))

gg_scatter_slope <- function(x, y, dt){
  ggplot(data = dt) + 
    geom_point(aes(UQ(rlang::sym(x)), UQ(rlang::sym(y))))
}

# dt[22,]
# which(dt$inlake_slope.y > 0.4)
# which(dt$inlake_slope.x > 0.4)
dt <- dplyr::filter(dt, inlake_slope.y < 0.4)

cowplot::plot_grid(
  gg_scatter_slope("slope_mean", "maxdepth", dt),
  gg_scatter_slope("inlake_slope.y", "maxdepth", dt), 
  gg_scatter_slope("slope_mean", "inlake_slope.y", dt)
  )
```

```{r buffer-vs-inlake_slope, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4}
lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>% 
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::arrange(desc(factor(shape_class, 
                        levels = c("neither", "concave", "convex"))))

# GGally::ggpairs(dplyr::select(test, matches("class")))

gg_slope_shape <- ggplot(data = test) + 
  geom_point(aes(slope_mean, inlake_slope.y, 
                             color = shape_class), size = 1) +
  geom_smooth(aes(slope_mean, inlake_slope.y, 
                             color = shape_class), data = dplyr::filter(test, shape_class != "neither"), 
              method = "lm", se = FALSE) +
  geom_segment(aes(x = 0, xend = 0.35, y = 0, yend = 0.35), 
               linetype = "dashed", size = 0.4) +
  # geom_abline(slope = 1, intercept = 0) +
  xlab("Nearshore slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "")

gg_slope_reservoir <- ggplot(data = dplyr::filter(test, !is.na(reservoir_class))) +
  geom_point(aes(slope_mean, inlake_slope.y, 
                                 color = reservoir_class), size = 1) +
  geom_smooth(aes(slope_mean, inlake_slope.y, 
                                 color = reservoir_class), 
              method = "lm", se = FALSE) +
  geom_segment(aes(x = 0, xend = 0.35, y = 0, yend = 0.35), 
               linetype = "dashed", size = 0.4) +
  xlab("Nearshore slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "")
```

```{r 01_geometry_base, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4.8, fig.height=2.3}
theme_opts <- cowplot::theme_cowplot() + 
  theme(legend.position = "none", 
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 11))
text_size <- 3

# distance data
dt         <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
hypso_pred <- read.csv("../data/00_hypso/hypso_predictors.csv", 
                       stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
dt <- left_join(dt, dplyr::select(hypso_pred, lagoslakeid, reservoir_class, shape_class), 
                                  by = c("llid" = "lagoslakeid"))

gg_dist_base <- st_drop_geometry(dt) %>%
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::filter(dist_viscenter >= dist_deepest) %>%
  dplyr::filter(dist_deepest < 3200) %>%
  ggplot(aes(x = dist_deepest, y = dist_viscenter)) + 
  geom_point(size = 1) +
  geom_smooth(method = "lm", se = FALSE) +
  annotate("text", x = 2550, y = 300, label = "R^2 == 0.80",
                                parse = TRUE, size = text_size) +
  annotate("text", x = 2100, y = 150, 
           label = get_formula(
             lm(dist_viscenter ~ dist_deepest, data = st_drop_geometry(dt))), 
           size = text_size) +
  labs(color = "") +
  xlab("Deepest point distance (m)") +
  ylab("Distance to \n lake center (m)") +
  xlim(0, 3000) +
  theme_opts

# slope data
bp <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
  mutate(llid = as.character(llid))
dt <- left_join(dt, st_drop_geometry(bp), 
                by = "llid") %>%
  dplyr::filter(!is.na(inlake_slope_mean)) %>%
  dplyr::filter(inlake_slope.y < 0.4)
lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>% 
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::arrange(desc(factor(shape_class, 
                        levels = c("neither", "concave", "convex"))))

gg_slope_base <- ggplot(data = test) + 
  geom_point(aes(slope_mean, inlake_slope.y), size = 1) +
  geom_smooth(aes(slope_mean, inlake_slope.y), method = "lm", se = FALSE) +
  geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
  alpha = 1, fill = "white") +
  annotate("text", x = 0.34, y = 0.03, label = "R^2 == 0.17",
                                parse = TRUE, size = text_size) +
  annotate("text", x = 0.34, y = 0.01, 
           label = get_formula(
             lm(inlake_slope.y ~ slope_mean, data = test)), 
           size = text_size) +
  # geom_abline(slope = 1, intercept = 0) +
  xlab("Nearshore slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "") +
  theme_opts

cowplot::plot_grid(
  gg_dist_base, 
  gg_slope_base, 
  rel_widths = c(1, 0.9)
)

```

```{r 01_geometry_grid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5.5, fig.height=4.2}
theme_opts <- cowplot::theme_cowplot() + 
  theme(legend.position = "none", 
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 11))
gg_spacer <- ggplot() + geom_blank() + theme(panel.background = element_rect(fill = NA), plot.background = element_rect(fill = NA))

legend_shape     <- cowplot::get_legend(gg_scatter_shape +
                                          theme(legend.direction = "horizontal", legend.background = element_rect(fill = NA)))
legend_reservoir <- cowplot::get_legend(gg_scatter_reservoir + labs(color = "") +
                                          theme(legend.direction = "horizontal", legend.background = element_rect(fill = NA)))
legend_panel     <- cowplot::plot_grid(gg_spacer, legend_shape, 
                                       gg_spacer, legend_reservoir, 
                                       nrow = 1, rel_widths = c(0.2, 1, 0.08, 1),
                                       align = "hv")

text_size <- 2.5
cowplot::plot_grid(
  cowplot::plot_grid(
  cowplot::plot_grid(gg_scatter_shape + 
                       labs(color = "") +
                       # theme(axis.title.x = element_blank()) +
                       annotate("text", x = 2250, y = 300, label = "R^2 == 0.80",
                                parse = TRUE, size = text_size) +
                       theme_opts,
                     gg_spacer,
                     gg_scatter_reservoir + 
                       labs(color = "") +
                       # annotate("text", x = 1800, y = 500, label = "R^2 == 0.80",
                       #          parse = TRUE) +
                       theme(#axis.title.x = element_blank(), 
                             axis.title.y = element_blank()) +
                       theme_opts, 
                     ncol = 3, rel_widths = c(1, 0, 1)),
  cowplot::plot_grid(ggplot() + geom_blank() + 
                       theme_opts + theme(axis.line = element_blank()), 
                     gg_slope_shape +
                       annotate("text", x = 0.36, y = 0.03, label = "R^2 == 0.17",
                                parse = TRUE, size = text_size) +
                       theme_opts,
                     gg_spacer,
                     gg_slope_reservoir + 
                         # annotate("text", x = 0.3, y = 0.03, label = "R^2 == 0.17",
                         #        parse = TRUE) +
                       theme(axis.title.y = element_blank()) +
                       theme_opts,
                     ncol = 4, rel_widths = c(0.11, 1, 0.11, 1)), 
                     rel_widths = c(0.7, 1), nrow = 2), 
 legend_panel, rel_heights = c(1, 0.12), ncol = 1
)
```

```{r slope_by_area_depth, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=5.5, fig.height=4.2}

nearshore     <- read.csv("../data/00_geometry/nearshore.csv",
                      stringsAsFactors = FALSE) %>%
                  dplyr::select(-inlake_slope)
dt_raw        <- read.csv("../data/lagosus_depth_predictors.csv", 
                      stringsAsFactors = FALSE) %>%
                  left_join(nearshore, b = c("lagoslakeid" = "llid")) %>% 
                  dplyr::filter(shape_class %in% c("concave", "convex")) %>% 
                  dplyr::filter(inlake_slope < 0.6)

breaks_dt <- classInt::classIntervals(dt_raw$lake_waterarea_ha, n = 4)$brks
dt        <- dt_raw %>%
  dplyr::mutate(x_bins = cut(dt_raw$lake_waterarea_ha,
      breaks = breaks_dt)) %>%
  dplyr::filter(!is.na(x_bins))
dt <- dt %>%
  dplyr::mutate(x_bins = forcats::fct_recode(x_bins, 
        !!!setNames(levels(dt$x_bins), 
                    paste0("< ", as.character(round(breaks_dt[-1], 1))))
      )
    )

gg_area <- ggplot(data = dt) +
   geom_point(aes(x = slope_mean, y = inlake_slope, color = x_bins)) +
   geom_smooth(aes(slope_mean, inlake_slope, color = x_bins),
               method = "lm", se = FALSE) + 
  facet_wrap(~shape_class) +
    ggtitle("Area (ha)")

breaks_dt <- classInt::classIntervals(dt_raw$lake_maxdepth_m, n = 4)$brks
dt        <- dt_raw %>%
  dplyr::mutate(x_bins = cut(dt_raw$lake_maxdepth_m,
      breaks = breaks_dt)) %>%
  dplyr::filter(!is.na(x_bins))
dt <- dt %>%
  dplyr::mutate(x_bins = forcats::fct_recode(x_bins, 
        !!!setNames(levels(dt$x_bins), 
                    paste0("< ", as.character(round(breaks_dt[-1], 1))))
      )
    )

gg_depth <- ggplot(data = dt) +
   geom_point(aes(x = slope_mean, y = inlake_slope, color = x_bins)) +
   geom_smooth(aes(slope_mean, inlake_slope, color = x_bins),
               method = "lm", se = FALSE) +
    facet_wrap(~shape_class) +
    ggtitle("Max Depth (m)")

theme_opts <- cowplot::theme_cowplot() +
  theme(legend.position = "bottom", 
        legend.title = element_blank(), 
        panel.spacing = unit(1, "lines")) 
cowplot::plot_grid(
  gg_area + theme_opts +
    guides(color = guide_legend(nrow = 2, byrow = TRUE)),
  gg_depth + theme_opts + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank(), 
          axis.line.y = element_blank(),
          axis.ticks.y = element_blank()) +
    guides(color = guide_legend(nrow = 2, byrow = TRUE)),
  rel_widths = c(1, 0.87))

```