### 2020-11-23 figures/01_contrasts.Rmd
more name change propagation

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r define-plots, echo=FALSE}
contrast_plot <- function(..., plot_type){
  theme_opts <- cowplot::theme_cowplot()

  if(plot_type == "violin"){
    violin_plot(...) + theme_opts
  }else{
    pointrange_plot(...) + theme_opts
  }
}

violin_plot <- function(x, grp, varb, label){
  ggplot(data = x) +
  geom_violin(aes(x = UQ(rlang::sym(grp)), y = UQ(rlang::sym(varb))), 
              adjust = 0.8, scale = "count", 
              draw_quantiles = c(0.25, 0.5, 0.75)) +
    ylab(label) + xlab("")
}

quantile_collapse <- function(x){
  list(tibble::enframe(
    quantile(x, probs = c(0.05,0.5,0.95), na.rm = TRUE)))
}

pointrange_plot <- function(x, grp, varb, label){
  x %>%
  group_by(UQ(rlang::sym(grp))) %>%
    dplyr::select(UQ(rlang::sym(grp)), UQ(rlang::sym(varb))) %>%
    summarize_all(quantile_collapse) %>%
    unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = UQ(rlang::sym(grp)), y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab(label) + xlab("")
}
```

```{r gg_depth_shape_class, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}

dt <- read.csv("../data/00_hypso/hypso_predictors.csv", 
               stringsAsFactors = FALSE)

# gg_depth_shape_class <- dt %>%
#   dplyr::filter(shape_class != "neither") %>%  
#   contrast_plot(grp = "shape_class", varb = "lake_maxdepth_m", 
#                 label = "Depth (m)", plot_type = "violin")

gg_depth_shape_class <- dt %>%
  dplyr::filter(shape_class != "neither") %>%  
  contrast_plot(grp = "shape_class", varb = "lake_maxdepth_m", 
                label = "Depth (m)", plot_type = "pointrange")
```

```{r gg_area_shape_class, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_area_shape_class <- dt %>%
  dplyr::filter(shape_class != "neither") %>%  
  contrast_plot(grp = "shape_class", varb = "lake_waterarea_ha", 
                label = "Area (ha)", plot_type = "pointrange")
```

```{r gg_slope_shape_class, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_slope_shape_class <- dt %>%
  dplyr::filter(shape_class != "neither") %>%  
  contrast_plot(grp = "shape_class", varb = "inlake_slope_pnt", 
                label = "In-lake slope (m/m)", plot_type = "pointrange")
```

```{r gg_dist_shape_class, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_dist_shape_class <- dt %>%
  dplyr::filter(shape_class != "neither") %>%  
  contrast_plot(grp = "shape_class", varb = "dist_deepest", 
                label = "Distance to deepest point (m)", plot_type = "pointrange")
```

```{r gg_depth_connectivity, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_depth_conny_class <- dt %>%
  dplyr::filter(!is.na(lake_connectivity_class)) %>%
  group_by(lake_connectivity_class) %>%
  summarize(lake_maxdepth_m = list(tibble::enframe(
    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = lake_connectivity_class, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab("")
```

```{r gg_depth_glaciated, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}

gg_depth_glaciated_class <- dt %>%
  dplyr::filter(!is.na(lake_glaciatedlatewisc)) %>%
  group_by(lake_glaciatedlatewisc) %>%
  summarize(lake_maxdepth_m = list(tibble::enframe(
    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = lake_glaciatedlatewisc, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab("")
```

```{r gg_depth_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_depth_reservoir <- dt %>%
  dplyr::filter(!is.na(reservoir_class)) %>%  
  contrast_plot(grp = "reservoir_class", varb = "lake_maxdepth_m", 
                label = "Depth (m)", plot_type = "pointrange")
```

```{r gg_area_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_area_reservoir <- dt %>%
  dplyr::filter(!is.na(reservoir_class)) %>%  
  contrast_plot(grp = "reservoir_class", varb = "lake_waterarea_ha", 
                label = "Area (ha)", plot_type = "pointrange")
```

```{r gg_slope_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_slope_reservoir <- dt %>%
  dplyr::filter(!is.na(reservoir_class)) %>%  
  contrast_plot(grp = "reservoir_class", varb = "inlake_slope_pnt", 
                label = "In-lake slope (m/m)", plot_type = "pointrange")
```

```{r gg_dist_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_dist_reservoir <- dt %>%
  dplyr::filter(!is.na(reservoir_class)) %>%  
  contrast_plot(grp = "reservoir_class", varb = "dist_deepest", 
                label = "Distance to deepest point (m)", plot_type = "pointrange")
```

```{r gg_depth_xtabs, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
dt_sub <- dplyr::filter(dt, 
                        !is.na(shape_class), shape_class != "neither", 
                        !is.na(reservoir_class))
# group_by(dt_sub, lake_glaciatedlatewisc, reservoir_class) %>%
#   tally() %>%
#   arrange(lake_glaciatedlatewisc) %>%
#   gridExtra::grid.table()

(gg_depth_xtabs <- group_by(dt_sub, 
                           lake_glaciatedlatewisc, reservoir_class) %>%
    summarize(lake_maxdepth_m = list(tibble::enframe(
    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  tidyr::unite(cat, lake_glaciatedlatewisc:reservoir_class) %>%
  ggplot() +
geom_pointrange(aes(x = cat, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab(""))

```

```{r gg_shorelinedev_shape_class, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# compare shoreline dev for each shapeclass
gg_shorelinedev_shape_class <- dt %>%
  dplyr::filter(shape_class != "neither") %>%  
  contrast_plot(grp = "shape_class", varb = "lake_shorelinedevfactor", 
                label = "Shoreline Dev.", plot_type = "pointrange")
```

```{r 01_contrasts, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 3.5}
theme_opts <- theme(axis.text.x = element_text(angle = 90), 
                    axis.title = element_blank(),
                    plot.title = element_text(size = 8))

gg_shape <- cowplot::plot_grid(
  gg_depth_shape_class + ggtitle("Depth (m)") + theme_opts,
  gg_area_shape_class + ggtitle("Area (ha)") + theme_opts, 
  gg_slope_shape_class + ggtitle("In-lake slope (m/m)") + theme_opts, 
  gg_dist_shape_class + ggtitle("Dist. deepest pnt (m)") + theme_opts, 
  nrow = 1)

gg_reservoir <- cowplot::plot_grid(gg_depth_reservoir + theme_opts, 
                                   gg_area_reservoir + theme_opts, 
                                   gg_slope_reservoir + theme_opts, 
                                   gg_dist_reservoir + theme_opts, 
                                   nrow = 1)

cowplot::plot_grid(gg_shape, gg_reservoir, 
                   nrow = 2, rel_heights = c(1, 0.76))

cowplot::plot_grid(gg_depth_xtabs + theme_opts, 
                   ggplot() + geom_blank() + 
                     theme(panel.background = element_blank()), nrow = 1)
```

```{r gg_tally_profileshapeclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# panel histograms that tally number of lakes with max depth by profileshape_class
# include sum across all types

gg_tally_profileshapeclass <- read.csv(
  "../data/00_hypso/hypso_predictors.csv", stringsAsFactors = FALSE) %>% 
  group_by(shape_class) %>%
  tally() %>%
  ggplot() + 
  geom_col(aes(x = shape_class, y = n))
```

```{r gg_tally_programtype, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide'}
# tally number of lakes with max depth by programtype_depth
gg_tally_programtype <- read.csv(
  "../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>% 
   dplyr::filter(!is.na(lake_maxdepth_m)) %>%
   group_by(sourcetype_depth) %>% 
   tally() %>% 
   ggplot(aes(x = sourcetype_depth, y = n)) + 
   geom_col() + 
   geom_text(aes(label = n), vjust = -0.5) + 
   ylim(0, 13000) +
  xlab("Source type")
```

```{r gg_tally_maxdepthclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# panel histograms that tally number of lakes with max depth by maxdepth_class
# include sum across all types

dt <- read.csv(
  "../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE)
max_depth <- dt$lake_maxdepth_m
gg_tally_maxdepthclass <- dt %>%
  mutate(maxdepth_class = 
           smart_cut(lake_maxdepth_m, c(10, 30, 70, 200, max(max_depth)), 
                     brackets = NULL, sep = " - ")
         ) %>%
  group_by(maxdepth_class) %>%
  tally() %>%
  ggplot(aes(x = maxdepth_class, y = n)) +
  geom_col() + 
  geom_text(aes(label = n), vjust = -0.5) +
  ylim(0, 15000) + 
  xlab("Max depth class (m)")
```

```{r gg_tally_waterareaclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# panel histograms that tally number of lakes with max depth by waterarea_class
# include sum across all types
gg_tally_waterareaclass <- read.csv("../data/lagosus_depth_predictors.csv", 
               stringsAsFactors = FALSE) %>%
  mutate(area_class =
           smart_cut(lake_waterarea_ha, c(1, 4, 40, 80, 300, 400,
                                               1000, 20000, Inf),
             labels = ~paste(sep="-", thousand_k(.y[1]), thousand_k(.y[2])))
         ) %>%
  group_by(area_class) %>%
  tally() %>%
  ggplot(aes(x = area_class, y = n)) +
  geom_col() + 
  geom_text(aes(label = n), vjust = -0.5) +
  ylim(0, 11000) + 
  xlab("Lake area class (ha)")
```

```{r gg_tally_hypsoclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_tally_hypsoclass <- read.csv(
  "../data/lagosus_depth_predictors.csv", 
               stringsAsFactors = FALSE) %>%
  dplyr::filter(!is.na(shape_class)) %>%
  distinct(lagoslakeid, shape_class, .keep_all = TRUE) %>%
  group_by(shape_class) %>%
  tally() %>%
  ggplot(aes(x = shape_class, y = n)) +
  geom_col() + 
  geom_text(aes(label = n), vjust = -0.5) +
  ylim(0, 4500) + 
  xlab("Shape class")
```

```{r 01_contrasts_tally, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 7.6}
cowplot::plot_grid(gg_tally_maxdepthclass, 
                   gg_tally_waterareaclass, gg_tally_hypsoclass, 
                   ncol = 1)
```

```{r 01_programs_tally, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide'}
# number of programs in each type
read.csv(
  "../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>% 
   dplyr::filter(!is.na(lake_maxdepth_m)) %>%
   distinct(sourcetype_depth, sourceurl_depth) %>% 
   ungroup() %>% group_by(sourcetype_depth) %>%
   tally() %>%
   ggplot(aes(x = sourcetype_depth, y = n)) + 
   geom_col() + 
   geom_text(aes(label = n), vjust = -0.5) + 
   ylim(0, 250) +
  xlab("Program type")
```

```{r gg_tally_glaciated_vs_shapeclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
dt_sub <- dplyr::filter(dt, !is.na(shape_class), shape_class != "neither")
as.data.frame(
  table(paste(dt_sub$lake_glaciatedlatewisc, dt_sub$shape_class))) %>%
  tidyr::separate(Var1, c("glaciation", "shape_class"), sep = " ") %>%
  ggplot() + 
  geom_point(aes(glaciation, shape_class, size = Freq)) +
  scale_size_continuous(range = c(5, 20)) +
  theme(legend.position = "none")
```

```{r gg_effort, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height = 3.7, fig.width=2.7}
# compare characteristics of bathy lakes versus overall lakes

# load LAGOSUS DEPTH
lg_raw <- lagosus_load("depth")$depth$depth
lg <- lg_raw %>%
  mutate(lagos_effort = 
           forcats::fct_collapse(lg_raw$lagos_effort, 
                                 bathymetry = "bathymetry", 
                                 other = c("LAGOSNE", "manual", "NLA")))

# library(tigris)
# urbanareas <- tigris::urban_areas()
# urbanareas <- st_as_sf(urbanareas)
# urbanareas <- st_transform(urbanareas, LAGOSUSgis:::albers_conic())
# urbanareas <- tidyr::separate(urbanareas, NAME10,
#                               c("urban_name", "urban_state"), sep = ", ")
# urbanareas <- dplyr::filter(urbanareas, !(urban_state %in% c("AK", "HI")))
# urbanareas <- urbanareas[st_area(urbanareas) > 
#                      units::as_units(9500000, "m2"),]
# for(i in seq_len(nrow(urbanareas))){
#   print(i)
#   urbanareas[i,] <- rmapshaper::ms_simplify(urbanareas[1,])
# }
# saveRDS(urbanareas, "../data/urbanareas.rds")
urbanareas <- readRDS("../data/urbanareas.rds")

lg$urban_distance <- apply(
  st_distance(
    st_transform(
      LAGOSUS::coordinatize(dt = lg, crs_in = 4326), 
      LAGOSUSgis:::albers_conic()),
    urbanareas), 
  1, function(x) x[which.min(x)])
lg$urban_distance <- lg$urban_distance * 0.001 # m to km

theme_opts <- theme(axis.title.x = element_blank())
cowplot::plot_grid(
  gg_quantdot(lg, lagos_effort, "lake_maxdepth_m") +
    ylab("Max depth (m) \n") +
    theme_opts +
    theme(axis.text.x = element_blank()), 
  gg_quantdot(lg, lagos_effort, "lake_waterarea_ha") +
    scale_y_log10() +
    ylab("Area (ha)") +
    theme_opts +
    theme(axis.text.x = element_blank()), 
  gg_quantdot(lg, lagos_effort, "urban_distance") +
    scale_y_log10() +
    ylab("Distance to urban (km)") +
    theme_opts,
  rel_heights = c(0.85, 0.85, 1),  ncol = 1)
```