### 2020-11-23 manuscript/figures.Rmd
rm predictor table to sync with ms

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{colortbl}
- \usepackage{float}
- \usepackage{caption}
---

\arrayrulecolor{white}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", dev = c("pdf"))
knitr::opts_chunk$set(fig.pos = 'H')
```

\pagebreak

<!-- \captionsetup[figure]{name=Table} -->
<!-- \captionsetup[figure]{labelformat=empty} -->
<!-- ```{r 01_predictors, fig.cap = "Table 1: Summary of lake characteristics for the present study (and for lakes in the contiguous United States from <LAGOSUS-Depth citation>). Predictor variables for computing random forest offsets (Eq 2) are printed in bold face. Dashes (-) indicate an identical sample size among this study and that of the contiguous United States.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'} -->
<!-- knitr::include_graphics("../tables/01_predictors.pdf") -->
<!-- ``` -->

<!-- \captionsetup[figure]{name=Figure} -->
<!-- \captionsetup[figure]{labelformat=default} -->
<!-- \setcounter{figure}{0} -->

![Diagram showing the relations between true (black) and proxy (orange) metrics of lake geometry. Geometric depth calculated via Equation 1 requires a single distance and slope metric.](../figures/slope_diagram_new.pdf){width=75%}

```{r 01_geometry_base, fig.cap = "Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope. A best-fit line and coefficient of determination is shown to illustrate representativeness.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
knitr::include_graphics("../figures/01_geometry_base-1.pdf")
```

\captionsetup[figure]{name=Table}
\captionsetup[figure]{labelformat=empty}
```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuracy metrics (RMSE = root mean square error, R2 = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
# see tables/02_model_metrics.Rmd
knitr::include_graphics("../tables/02_model_metrics.pdf")
```
\captionsetup[figure]{name=Figure}
\captionsetup[figure]{labelformat=default}
\setcounter{figure}{3}

```{r depth_model_grid_resid, fig.cap = "Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
# see 02_depth_model.Rmd
knitr::include_graphics("../figures/02_depth_model_grid_resid-1.pdf")
```