### 2020-12-07 figures/02_depth_model.Rmd
add xaxis labels to residual density plot

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
if(interactive()){
  setwd("figures")
}
source("../scripts/99_utils.R")
```

```{r depth_model_metrics, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=2.6, fig.width=4}
dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds") %>%
  mutate_if(is.numeric, function(x) as.character(round(x, 2))) %>%
  mutate(rsq = as.numeric(rsq)) %>%
  mutate(rsq = case_when(
    as.character(model) == "maxdepth_true_true" ~ NA_real_,
    TRUE ~ rsq)) %>%
  mutate(rmse = case_when(
    as.character(model) == "maxdepth_true_true" ~ " ",
    TRUE ~ paste0(rmse, " m"))) %>%
  mutate(model = gsub("maxdepth_", "", model)) %>%
  mutate(model = gsub("false", "proxy", model)) %>%
  tidyr::separate(model, c("slope", "distance")) %>%
  dplyr::select(slope, distance, everything(), -.estimator)

ggplot(data = dt_metrics, 
                    aes(y = slope, 
                        x = reorder(distance, desc(distance)),
                        fill = rsq)) +
  geom_tile(color = "black") +
  scale_fill_viridis_c(begin = 0.4, end = 0.7, na.value = "transparent") +
  scale_x_discrete(limits = rev(levels(dt_metrics$slope)), 
                   position = "top") +
  geom_text(aes(label = rmse), vjust = .5, fontface  = "bold", alpha = 1) +
  xlab("distance") + ylab("slope") +
  cowplot::theme_cowplot() +
  theme(axis.line = element_blank(), 
        axis.ticks = element_blank())
```

```{r 02_depth_model_importance, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.2, fig.width=5}
dt_fits      <- readRDS("../data/01_depth_model/depth_fits.rds")
taxonomy_key <- read.csv("../data/taxonomy.csv", stringsAsFactors = FALSE)

tidy_importance <- function(x){
  # x <- dt_fits[[2]]$fit$fit
  res <- ranger::importance(x) %>%
  data.frame() %>% cbind(row.names(.), .[,1]) %>%
  setNames(c("percent_inc_mse", "variable"))
  row.names(res) <- NULL
  res <- res[,c(2, 1)]
  res <- arrange(res, percent_inc_mse)
  # res$variable <- factor(res$variable, levels = res$variable)
  res <- mutate(res, variable = as.character(variable))
  res <- mutate(res, variable = case_when(
    stringr::str_detect(variable, "maxdepth") ~ "trig_depth", 
    TRUE ~ variable)
    )
  res
}

dt_imp <- lapply(2:4, function(i) tidy_importance(dt_fits[[i]]$fit$fit))
model_names <- c("true_proxy", 
                 "proxy_true", 
                 "proxy_proxy")
dt_imp <- bind_rows(lapply(1:3, function(i){
  mutate(dt_imp[[i]], model = model_names[i])
  }))
dt_imp$model <- factor(dt_imp$model, levels = model_names)

var_order <- 
  left_join(dt_imp, taxonomy_key, by = c("variable" = "col_name")) %>%
  group_by(pretty) %>% 
  summarize(max = max(percent_inc_mse)) %>%
  arrange(max) %>%
  pull(pretty)
dt_imp <- left_join(dt_imp, taxonomy_key, by = c("variable" = "col_name")) %>%
  mutate(pretty = factor(pretty, levels = var_order)) %>%
  mutate(variable = pretty) %>%  dplyr::select(-description) %>%
  dplyr::select(-pretty)

ggplot() + 
  geom_point(data = dt_imp, aes(percent_inc_mse, variable, color = model)) +   
  labs(color = "slope-distance\nmeasure") +
  geom_hline(linetype = 3, size = 1, colour = "gray",
                 yintercept = 1:nlevels(dt_imp$variable)) +
  theme(text = element_text(family="Times"),
        panel.background = element_blank(), panel.grid = element_blank(),
        panel.border = element_rect(fill = NA), # legend.position = "none", 
        axis.title.x = element_text(family = "sans", face = "bold", 
        vjust = -0.5, size = 10), 
        axis.text = element_text(family="sans",size = 9, 
                                 color = "black"), 
         legend.key = element_blank()) + 
        ylab("") + xlab("Percent Increase in Mean Square Error")
```

```{r 02_depth_model_grid_resid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5, fig.width=8}

lgus <- read.csv("../data/lagosus_depth_predictors.csv", 
                 stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
  dplyr::select(-lake_maxdepth_m) %>%
  distinct(lagoslakeid, .keep_all = TRUE)
dt   <- readRDS("../data/01_depth_model/depth_grid.rds") %>% # 05_depth_model.R
  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
  mutate(resid = case_when(maxdepth == "maxdepth_true_true" ~ NA_real_, 
                           TRUE ~ resid)) %>%
  mutate(maxdepth = factor(maxdepth, levels = c("maxdepth_true_true", "maxdepth_true_false", 
                                  "maxdepth_false_true", "maxdepth_false_false"))) %>%
  left_join(lgus, by = "lagoslakeid")

gg_shape <- lapply(unique(dt$maxdepth), function(x){ 
  ggplot(data = dplyr::filter(dt, maxdepth == x)) + 
  geom_density(aes(x = resid, fill = shape_class), alpha = 0.75, adjust = 1.3) +
  scale_fill_manual(values = hue_pal()(3)[1:2]) +
  geom_vline(aes(xintercept = 0), linetype = "dashed") +
    xlim(-25, 25) +
    xlab("Residual") +
    theme_cowplot() +
    theme(legend.position = "bottom",
        strip.background = element_blank(), 
        axis.text = element_text(size = 9.5),
        strip.text = element_blank(),
        plot.margin = unit(c(1.2,0.2,0.2,0.2), "cm"))  +
  guides(fill = guide_legend(nrow = 2))
  })
gg_shape[[1]] <- ggplot() + geom_blank() + cowplot::theme_cowplot() +
  theme(axis.line = element_blank())
leg            <- cowplot::get_legend(gg_shape[[2]] + labs(fill = "Shape class") +
                                        theme(legend.direction = "horizontal") +
                                        guides(fill = guide_legend(ncol = 2)))
theme_opts     <- function(x) x + theme(legend.position = "none", 
                                        axis.title = element_blank())
rm_legend     <- function(x) x + theme(legend.position = "none")
gg_shape      <- lapply(gg_shape, rm_legend)
gg_shape[[2]] <- gg_shape[[2]] + theme(axis.text = element_blank(), 
                                       axis.title = element_blank())
gg_shape[[3]] <- gg_shape[[3]] + theme(axis.title = element_blank())
gg_shape[[4]] <- gg_shape[[4]] + theme(axis.text.y = element_blank(), 
                                       axis.title = element_blank())

gg_shape <- cowplot::plot_grid(gg_shape[[1]], 
          cowplot::plot_grid(
            cowplot::plot_grid(gg_shape[[1]], gg_shape[[2]], 
                               rel_widths = c(1, 0.85)),
            cowplot::plot_grid(gg_shape[[3]], gg_shape[[4]], 
                               rel_widths = c(1, 0.85)), 
            ggdraw() + draw_label("Residual (m)", size = 10, y = 0.6),
            leg, 
            nrow = 4, rel_heights = c(0.85, 1, 0.1, 0.15)),

          ncol = 2, rel_widths = c(0.1, 1)) %>%
  ggdraw() +
  draw_label("proxy", x = 0.06, y = 0.32, angle = 90, size = 10) +
  draw_label("true", x = 0.06, y = 0.75, angle = 90, size = 10) +
  draw_label("slope", x = 0.01, y = 0.55, size = 10, angle = 90) +
  draw_label("proxy", x = 0.8, y = 0.9, size = 10) +
  draw_label("true", x = 0.35, y = 0.9, size = 10) +
  draw_label("distance", x = 0.55, y = 0.96, size = 10)#  +
  # draw_label("Residual distribution", 
  #            x = 0.55, y = 0.99, size = 10, fontface = "bold")

gg_reservoir <- lapply(unique(dt$maxdepth), function(x){ 
  ggplot(data = dplyr::filter(dt, maxdepth == x)) + 
  geom_density(aes(x = resid, fill = reservoir_class), alpha = 0.75, adjust = 1.3) +
  geom_vline(aes(xintercept = 0), linetype = "dashed") +
    xlim(-25, 25) +
    xlab("Residual") +
    theme_cowplot() +
    theme(legend.position = "bottom",
        strip.background = element_blank(), 
        strip.text = element_blank(),
        axis.text = element_text(size = 9.5),
        plot.margin = unit(c(1.2,0.2,0.2,0.2), "cm"))  +
  guides(fill = guide_legend(nrow = 2))
  })
gg_reservoir[[1]] <- ggplot() + geom_blank() + cowplot::theme_cowplot() +
  theme(axis.line = element_blank())
leg            <- cowplot::get_legend(gg_reservoir[[2]] + labs(fill = "Reservoir class") +
                                        theme(legend.direction = "horizontal") +
                                        guides(fill = guide_legend(ncol = 2)))
theme_opts     <- function(x) x + theme(legend.position = "none", 
                                        axis.title = element_blank())
rm_legend     <- function(x) x + theme(legend.position = "none")
gg_reservoir      <- lapply(gg_reservoir, rm_legend)
gg_reservoir[[2]] <- gg_reservoir[[2]] + theme(axis.text = element_blank(), 
                                       axis.title = element_blank())
gg_reservoir[[3]] <- gg_reservoir[[3]] + theme(axis.title = element_blank())
gg_reservoir[[4]] <- gg_reservoir[[4]] + theme(axis.text.y = element_blank(), 
                                       axis.title = element_blank())

gg_reservoir <- cowplot::plot_grid(gg_reservoir[[1]], 
          cowplot::plot_grid(
            cowplot::plot_grid(gg_reservoir[[1]], gg_reservoir[[2]], 
                               rel_widths = c(1, 0.85)),
            cowplot::plot_grid(gg_reservoir[[3]], gg_reservoir[[4]], 
                               rel_widths = c(1, 0.85)), 
            ggdraw() + draw_label("Residual (m)", size = 10, y = 0.6),
            leg, 
            nrow = 4, rel_heights = c(0.85, 1, 0.1, 0.15)), 
          ncol = 2, rel_widths = c(0.1, 1)) %>%
  ggdraw() +
  # draw_label("proxy", x = 0.06, y = 0.32, angle = 90, size = 10) +
  # draw_label("true", x = 0.06, y = 0.75, angle = 90, size = 10) +
  # draw_label("slope", x = 0.01, y = 0.55, size = 10, angle = 90) +
  draw_label("proxy", x = 0.8, y = 0.9, size = 10) +
  draw_label("true", x = 0.35, y = 0.9, size = 10) +
  draw_label("distance", x = 0.55, y = 0.96, size = 10) # +
  # draw_label("Residual distribution", 
  #            x = 0.55, y = 0.99, size = 10, fontface = "bold")

cowplot::plot_grid(gg_shape, gg_reservoir, ncol = 2, rel_widths = c(1, 0.9))
```

```{r 02_depth_model_fitted, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
dt <- readRDS("../data/01_depth_model/depth_model.rds")

gg_noshape <- ggplot() +
  geom_point(data = dplyr::filter(dt$no_shape, shape_class == "convex"),
             aes(y = lake_maxdepth_m, x = fitted_values, color = shape_class),
             alpha = 0.3) +
  geom_point(data = dplyr::filter(dt$no_shape, shape_class == "concave"),
             aes(y = lake_maxdepth_m, x = fitted_values, color = shape_class),
             alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0) +
  ylab("Measured max depth") + xlab("Predicted max depth") +
  ylim(0, 80) + xlim(0, 80) +
  labs(color = "") +
  ggtitle("Max depth ~ Lake Area + Buffer slope")
  # theme(legend.position = "none")

gg_shape <- ggplot() +
  geom_point(data = dplyr::filter(dt$shape, shape_class == "convex"),
             aes(y = lake_maxdepth_m, x = fitted_values, color = shape_class),
             alpha = 0.3) +
  geom_point(data = dplyr::filter(dt$shape, shape_class == "concave"),
             aes(y = lake_maxdepth_m, x = fitted_values, color = shape_class),
             alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0) +
  ylab("Measured max depth") + xlab("Predicted max depth") +
  ylim(0, 80) + xlim(0, 80) +
  labs(color = "") +
  ggtitle("Max depth ~ Lake Area + Buffer slope + Shape class")

legend <- get_legend(
  # create some space to the left of the legend
  gg_noshape + theme(#legend.box.margin = margin(0, 0, 0, 12), 
                     legend.direction = "horizontal")
)

theme_opts <- theme(legend.position = "none", 
                    plot.title = element_text(size = 8))
cowplot::plot_grid(gg_noshape + theme_opts, 
                   gg_shape + theme_opts,
                   legend, 
                   rel_heights = c(1, 0.1)
                   )

```

```{r 02_depth_model_resid, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# residual = observed - predicted
# positive value = underprediction
# negative value = overprediction

ggplot() +
  geom_boxplot(data = dt$no_shape, aes(y = resid, x = shape_class),
               outlier.shape = NA) +
  ylim(-20, 20)
```

```{r 02_depth_oliver, echo=FALSE, message=FALSE, eval=FALSE, warning=FALSE, results='hide', fig.height=3.4}
# setwd("figures")
oliver2015 <- LAGOSNE::lagos_load_oliver_2015() %>%
  mutate(lake_maxdepth_m_oliver = zmaxpredict)
dt <- readRDS("../data/01_depth_model/oliver_model.rds")

rmse_group <- group_by(dt$test, hu4_zoneid) %>%
  summarize(r_rmse = yardstick::rmse_vec(
    exp(lake_maxdepth_m), 
    exp(lake_maxdepth_m_predicted)) / mean(lake_maxdepth_m))
hu4_gis <- st_read("../data/gis.gpkg", layer = "hu4s_focal_simple") %>%
# ggplot() + 
#   geom_sf(data = hu4_gis) +
#   geom_sf_text(data = hu4_gis, aes(label = hu4_zoneid))
  left_join(rmse_group) %>%
  dplyr::filter(!is.na(r_rmse) & r_rmse != Inf & r_rmse < 12)
ggplot() +
  geom_sf(data = hu4_gis, aes(fill = r_rmse))

## predicted vs obs
ggplot(data = dt$test) + 
  geom_point(aes(x = lake_maxdepth_m_predicted, 
                 y = lake_maxdepth_m)) +
  geom_abline(aes(slope = 1, intercept = 0))

## calculate metrics
gridExtra::grid.arrange(gridExtra::tableGrob(
bind_rows(yardstick::rmse(dt$test, 
                          exp(lake_maxdepth_m), exp(lake_maxdepth_m_predicted)), 
          yardstick::rsq(dt$test,
                         exp(lake_maxdepth_m),
                         exp(lake_maxdepth_m_predicted)))))

## replication vs oliver
# dt$test <- dt$test %>%
#   left_join(dplyr::select(dplyr::mutate(oliver2015, lagoslakeid = as.character(lagoslakeid)), lagoslakeid,
#                           lake_maxdepth_m_oliver))
# ggplot(data = dt$test) + 
#   geom_point(aes(x = exp(lake_maxdepth_m_predicted), 
#                  y = lake_maxdepth_m_oliver)) +
#   geom_abline(aes(slope = 1, intercept = 0))

cowplot::plot_grid(
  ggplot(data = oliver2015) +
    geom_point(aes(x = zmaxpredict, y = zmaxobs)) +
    geom_abline(aes(slope = 1, intercept = 0)), 
  ggplot(data = oliver2015) +
    geom_point(aes(x = log(zmaxpredict), y = log(zmaxobs))) +
    geom_abline(aes(slope = 1, intercept = 0))
)
```