### 2021-08-10 figures/00_maps.Rmd
linting

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
  fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 00_map, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
study_bbox    <- st_read("../data/gis.gpkg", "study_bbox", quiet = TRUE)
dt            <- st_read("../data/gis.gpkg", "dt", quiet = TRUE)
max_depth <- dt$lake_maxdepth_m
dt <- dt %>%
  mutate(lake_maxdepth_m = round(lake_maxdepth_m, 2)) %>%
  mutate(maxdepth_class =
    smart_cut(lake_maxdepth_m, c(10, 30, 70, 200, max(max_depth)),
      brackets = NULL, sep = " - ")
  ) %>% # unique(dt$maxdepth_class)
  dplyr::filter(!is.na(maxdepth_class))
states_all    <- st_read("../data/gis.gpkg", "states_all", quiet = TRUE)
states_focal  <- st_read("../data/gis.gpkg", "states_focal", quiet = TRUE)
# hu4_focal     <- st_read("../data/gis.gpkg", "hu4s_focal_simple",
#   quiet = TRUE)
# hu4_focal     <- st_crop(st_transform(
#   lwgeom::st_make_valid(hu4_focal), st_crs(study_bbox)),
#   study_bbox)

dt_notsf <- cbind(st_drop_geometry(dt), st_coordinates(dt))

gg_states <-
  ggplot() +
  geom_sf(data = study_bbox[1, ], alpha = 1, size = 0.6,
    color = "black", fill = "#1D9DCF") +
  geom_sf(data = states_all, fill = "gray", color = "black", size = 0.3) +
  # geom_sf(data = outline, color = "black", alpha = 0, size = 0.8) +
  # geom_sf(data = hu4, fill = "grey") +
  geom_sf(data = states_focal, fill = "white", color = "black", size = 0.4) +
  geom_point_rast(data = dt_notsf, aes(x = X, y = Y, color = maxdepth_class),
    alpha = 0) +
  scale_color_manual(values = viridisLite::viridis(5)) +
  geom_point_rast(data = dplyr::filter(dt_notsf, lake_maxdepth_m < 10),
    aes(x = X, y = Y, color = maxdepth_class), size = 0.2, alpha = 0.8) +
  geom_point_rast(data = dplyr::filter(dt_notsf, lake_maxdepth_m > 10),
    aes(x = X, y = Y, color = maxdepth_class), size = 0.2, alpha = 0.8) +
  labs(color = "Max depth (m)") +
  coord_sf(datum =  NA) +
  theme_void()

gg_states
```

```{r 00_map_state, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# proportion of lakes > 4ha with max depth by state
lg <- lagosus_load("locus")
dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE)

us_states <- dplyr::filter(
  dplyr::select(states_all, postal, iso_a2), iso_a2 == "US")
dt_plot <- dt %>%
  distinct(lagoslakeid, .keep_all = TRUE) %>%
  dplyr::filter(lake_waterarea_ha >= 4) %>%
  group_by(lake_state) %>%
  summarize(prop_maxdepth = mean(!is.na(lake_maxdepth_m))) %>%
  arrange(desc(prop_maxdepth))
dt_plot <- dt_plot %>%
  left_join(us_states, by = c("lake_state" = "postal")) %>%
  dplyr::filter(!is.na(prop_maxdepth)) %>%
  distinct(lake_state, .keep_all = TRUE)
st_geometry(dt_plot) <- dt_plot$geom

ggplot() +
  geom_sf(data = dt_plot, aes(fill = prop_maxdepth)) +
  geom_sf_text(data = dt_plot,
    aes(label = round(prop_maxdepth, 2)),
    size = 2.5) +
  labs(fill = "") +
  coord_sf(datum = NA) + theme_void()

# limit to lakes with limno data
dt_plot <- dt %>%
  distinct(lagoslakeid, .keep_all = TRUE) %>%
  dplyr::filter(lake_waterarea_ha >= 4) %>%
  dplyr::filter(has_limno == 1) %>%
  group_by(lake_state) %>%
  summarize(prop_maxdepth = round(mean(!is.na(lake_maxdepth_m)), 2)) %>%
  arrange(desc(prop_maxdepth)) %>%
  left_join(dplyr::filter(
    dplyr::select(states_all, postal, iso_a2), iso_a2 == "US"),
  by = c("lake_state" = "postal")) %>%
  dplyr::filter(!is.na(prop_maxdepth)) %>%
  distinct(lake_state, .keep_all = TRUE)
st_geometry(dt_plot) <- dt_plot$geom

ggplot() +
  geom_sf(data = dt_plot, aes(fill = prop_maxdepth)) +
  geom_sf_text(data = dt_plot,
    aes(label = prop_maxdepth),
    size = 2.5) +
  labs(fill = "") +
  coord_sf(datum = NA) + theme_void()
```

```{r gg_bathy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
bathy_pnts <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
dt         <- st_read("../data/gis.gpkg", "dt", quiet = TRUE)
max_depth <- dt$lake_maxdepth_m
states_all    <- st_read("../data/gis.gpkg", "states_all", quiet = TRUE)
states_focal  <- st_read("../data/gis.gpkg", "states_focal", quiet = TRUE)
dt <- dt %>%
  mutate(lake_maxdepth_m = round(lake_maxdepth_m, 2)) %>%
  mutate(maxdepth_class =
    smart_cut(lake_maxdepth_m, c(10, 30, 70, 200, max(max_depth)),
      brackets = NULL, sep = " - ")
  ) %>% # unique(dt$maxdepth_class)
  dplyr::filter(!is.na(maxdepth_class))

dt_bathy <- st_drop_geometry(bathy_pnts) %>%
  left_join(
    dplyr::distinct(dplyr::mutate(dplyr::select(dt, lagoslakeid, maxdepth_class,
      has_limno, lake_waterarea_ha),
    lagoslakeid = as.character(lagoslakeid)), lagoslakeid, .keep_all = TRUE),
    by = c("llid" = "lagoslakeid"))

dt_bathy <- st_sf(dt_bathy, geometry = dt_bathy$geom)
dt_bathy <- mutate(dt_bathy,
  maxdepth_class = droplevels(dt_bathy$maxdepth_class))
dt_bathy <- dplyr::filter(dt_bathy, !is.na(maxdepth_class))

gg_bathy <-
  ggplot() +
  geom_sf(data = st_as_sfc(st_bbox(dt_bathy)), alpha = 1, size = 0.6,
    color = "black", fill = "#1D9DCF") +
  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))),
    fill = "gray", color = "black", size = 0.3) +
  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))),
    fill = "white", color = "black", size = 0.4) +
  geom_sf(data = dt_bathy, aes(color = maxdepth_class),
    size = 0.2, alpha = 0.8) +
  scale_color_manual(values = viridisLite::viridis(4)) +
  labs(color = "Max depth (m)  ") +
  coord_sf(xlim = c(-103.81483, -67.10548), ylim = c(37.5, 48.72102)) +
  theme_void()

gg_bathy
```

```{r gg_colocated, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
is_colocated <- read.csv("../data/lagosus_depth.csv",
  stringsAsFactors = FALSE) %>%
  dplyr::filter(
    lagos_effort %in% c("LAGOSNE", "bathymetry", "manual") &
      !is.na(lake_maxdepth_m) &
      # remove llids with duplicate manual depth)
      !(lagoslakeid %in% c(456169))) %>%
  dplyr::select(lagoslakeid, lake_state, lagos_effort, lake_maxdepth_m) %>%
  tidyr::pivot_wider(names_from = lagos_effort,
    values_from = c(lake_maxdepth_m)) %>%
  dplyr::filter(!is.na(bathymetry)) %>%
  mutate(combined = coalesce(manual, LAGOSNE)) %>%
  dplyr::filter(!is.na(combined)) %>%
  dplyr::filter(!is.na(bathymetry) & !is.na(combined)) %>%
  pull(lagoslakeid)

dt_bathy <- mutate(dt_bathy,
  colocated = case_when(llid %in% is_colocated ~ "colocated",
    TRUE ~ "bathymetry only")) %>%
  arrange(desc(colocated))

gg_colocated <- ggplot() +
  geom_sf(data = st_as_sfc(st_bbox(dt_bathy)), alpha = 1, size = 0.6,
    color = "black", fill = "#1D9DCF") +
  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))),
    fill = "gray", color = "black", size = 0.3) +
  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))),
    fill = "white", color = "black", size = 0.4) +
  geom_sf(data = dt_bathy, aes(color = colocated),
    size = 0.2, alpha = 0.8) +
  coord_sf(xlim = c(-103.81483, -67.10548), ylim = c(37.5, 48.72102)) +
  theme_void()

gg_colocated
```

```{r gg_shapeclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
dt           <- st_read("../data/gis.gpkg", "dt", quiet = TRUE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
  distinct(lagoslakeid, .keep_all = TRUE)
dt_withshape <- read.csv("../data/lagosus_depth_predictors.csv",
  stringsAsFactors = FALSE) %>%
  dplyr::distinct(lagoslakeid, shape_class) %>%
  dplyr::filter(!is.na(shape_class) & shape_class != "neither") %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
dt           <- inner_join(dt, dt_withshape) %>%
  arrange(desc(shape_class))

gg_shapeclass <- ggplot() +
  geom_sf(data = st_as_sfc(st_bbox(dt_bathy)), alpha = 1, size = 0.6,
    color = "black", fill = "#1D9DCF") +
  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))),
    fill = "gray", color = "black", size = 0.3) +
  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))),
    fill = "white", color = "black", size = 0.4) +
  geom_sf(data = dplyr::filter(dt, shape_class == "convex"),
    aes(color = shape_class), size = 0.03, alpha = 0.3) +
  geom_sf(data = dplyr::filter(dt, shape_class != "convex"),
    aes(color = shape_class), size = 0.07, alpha = 0.8) +
  labs(color = "Shape class     ") +
  coord_sf(xlim = c(-103.81483, -67.10548), ylim = c(37.5, 48.72102)) +
  theme_void()

gg_shapeclass

# table of shape_class by state
# st_drop_geometry(dt) %>%
#   group_by(lake_state, shape_class) %>%
#   count() %>%
#   tidyr::pivot_wider(names_from = shape_class, values_from = n)
```

```{r gg_reservoir, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
dt           <- st_read("../data/gis.gpkg", "dt", quiet = TRUE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
  distinct(lagoslakeid, .keep_all = TRUE)
dt_withres <- read.csv("../data/lagosus_depth_predictors.csv",
  stringsAsFactors = FALSE) %>%
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::distinct(lagoslakeid, reservoir_class) %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
dt           <- inner_join(dt, dt_withres) %>%
  arrange(desc(reservoir_class))

gg_reservoir <-
  ggplot() +
  geom_sf(data = st_as_sfc(st_bbox(dt_bathy)), alpha = 1, size = 0.6,
    color = "black", fill = "#1D9DCF") +
  geom_sf(data = st_crop(states_all, st_as_sfc(st_bbox(dt_bathy))),
    fill = "gray", color = "black", size = 0.3) +
  geom_sf(data = st_crop(states_focal, st_as_sfc(st_bbox(dt_bathy))),
    fill = "white", color = "black", size = 0.4) +
  geom_sf(data = dplyr::filter(dt, !is.na(reservoir_class)),
    aes(color = reservoir_class), size = 0.03, alpha = 0.3) +
  labs(color = "Reservoir class") +
  coord_sf(xlim = c(-103.81483, -67.10548), ylim = c(37.5, 48.72102)) +
  theme_void()

gg_reservoir
```

```{r 00_map_bathy, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=6.5}
theme_opts <- guides(color = guide_legend(
  override.aes = list(size = 2)))
cowplot::plot_grid(gg_bathy + theme_opts,
  gg_shapeclass + theme_opts,
  gg_reservoir + theme_opts,
  ncol = 1)
```