### 2021-11-10 figures/00_qa.Rmd
rm stale code

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 4, fig.height = 4)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 00_qa, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=2.5}
# variation among measurements for the same lake

dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
  group_by(lagoslakeid) %>%
  add_tally() %>%
  dplyr::filter(n > 1) %>%
  # dplyr::filter(lagoslakeid == 5426)
  summarize( # lake_maxdepth_std = sd(lake_maxdepth_m),
    lake_maxdepth_range = range(lake_maxdepth_m)[2] -
      range(lake_maxdepth_m)[1],
    lake_maxdepth_prop = lake_maxdepth_range / range(lake_maxdepth_m)[2]) %>%
  setNames(c("lagoslakeid",
    "Range", "Range / largest"))
# View(dplyr::arrange(dt, desc(Range)))
# plot(dt$Range, dt$`Range / largest`)

ggplot(data = tidyr::pivot_longer(dt, -lagoslakeid)) +
  geom_histogram(aes(x = value)) +
  facet_wrap(~name, scales = "free") +
  xlab("meters")
```

```{r lgnemanual-vs-bathy-depth, fig.height=2.2, fig.width=5}
dt <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE)

dt_sub <- dt %>%
  dplyr::filter(
    lagos_effort %in% c("LAGOSNE", "bathymetry", "manual") &
      !is.na(lake_meandepth_m)) %>%
  dplyr::select(lagoslakeid, lake_state, lagos_effort, lake_meandepth_m) %>%
  tidyr::pivot_wider(names_from = lagos_effort,
    values_from = c(lake_meandepth_m)) %>%
  dplyr::filter(!is.na(bathymetry)) %>%
  mutate(combined = coalesce(manual, LAGOSNE)) %>%
  dplyr::filter(!is.na(combined))

gg_mean <- ggplot() +
  geom_point(data = dt_sub,
    aes(x = bathymetry, y = combined, color = lake_state),
    alpha = 0.6) +
  geom_point(data = dplyr::filter(dt_sub, !(lake_state %in% c("ME"))),
    aes(x = bathymetry, y = combined, color = lake_state)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  labs(color = "") +
  xlab("Calculated (m)") + ylab("Reported (m)") +
  ggtitle("Mean Depth")

# plotly::ggplotly(gg)

dt_sub <- dt %>%
  dplyr::filter(
    lagos_effort %in% c("LAGOSNE", "bathymetry", "manual") &
      !is.na(lake_maxdepth_m) &
      !(lagoslakeid %in% c(456169))) %>% # remove llids with duplicate manual depth
  dplyr::select(lagoslakeid, lake_state, lagos_effort, lake_maxdepth_m) %>%
  tidyr::pivot_wider(names_from = lagos_effort,
    values_from = c(lake_maxdepth_m)) %>%
  dplyr::filter(!is.na(bathymetry)) %>%
  mutate(combined = coalesce(manual, LAGOSNE)) %>%
  dplyr::filter(!is.na(combined))

gg_max <- ggplot(data = dt_sub) +
  # geom_text(aes(x = bathymetry, y = combined, color = lake_state,
  #                label = lagoslakeid)) +
  geom_point(aes(x = bathymetry, y = combined, color = lake_state)) +
  geom_abline(aes(intercept = 0, slope = 1)) +
  xlab("Calculated (m)") + ylab("Reported (m)") +
  ggtitle("Max Depth")

cowplot::plot_grid(gg_max + cowplot::theme_cowplot() +
  theme(legend.position = "none"),
gg_mean + cowplot::theme_cowplot() +
  theme(axis.title.y = element_blank()),
rel_widths = c(0.8, 1))
```

```{r lgne-area_vs_lgus-area, eval=FALSE}
lg    <- lagosus_load("locus")
lg_ne <- lagosne_load()
dt    <- read.csv("../data/lagosus_depth.csv", stringsAsFactors = FALSE)

res <- dt %>%
  left_join(
    dplyr::select(lg$locus$locus_link, lagoslakeid, lagosne_lagoslakeid)) %>%
  left_join(
    dplyr::select(lg$locus$locus_characteristics,
      lagoslakeid, lake_totalarea_ha)) %>%
  dplyr::filter(!is.na(lagosne_lagoslakeid)) %>%
  dplyr::filter(lagoslakeid == lagosne_lagoslakeid) %>%
  dplyr::filter(!is.na(lake_maxdepth_m)) %>%
  distinct(lagoslakeid, .keep_all = TRUE) %>%
  dplyr::select(lagoslakeid, lagosne_lagoslakeid, lake_waterarea_ha,
    lake_totalarea_ha, lake_maxdepth_m) %>%
  left_join(dplyr::select(lg_ne$locus, lagosne_lagoslakeid = lagoslakeid,
    lake_area_ha)) %>%
  mutate(diff = lake_area_ha - lake_waterarea_ha) %>%
  arrange(desc(diff))

res2 <- res %>%
  dplyr::filter(abs(diff) < 0.1)
hist(res2$diff)

res3 <- res %>%
  dplyr::filter(abs(diff) > 1)
i <- 2
mapview(LAGOSUSgis::query_lake_poly(1222))
dplyr::filter(lg$locus$locus_characteristics, lagoslakeid == 14794)
dplyr::filter(dt, lagoslakeid == 14794)
dplyr::filter(lg_ne$locus, lagoslakeid == 14794)

+
  mapview(LAGOSNEgis::query_gis(
    "LAGOS_NE_All_Lakes_4ha", "lagoslakeid", res3$lagoslakeid[i]))

mapview(LAGOSNEgis::query_gis(
  "LAGOS_NE_All_Lakes_4ha", "lagoslakeid", 61727))
```