### 2021-11-10 README.Rmd
hide docker, link ms

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

[![Paper DOI](https://img.shields.io/badge/Paper-DOI-blue.svg)](https://doi.org) [![Code DOI](https://img.shields.io/badge/Code-DOI-blue.svg)](https://doi.org/) [![Data DOI](https://img.shields.io/badge/Data-10.6084/m9.figshare.12722246-blue.svg)](https://doi.org/10.6084/m9.figshare.12722246)

<!--- [![Docker Build](https://img.shields.io/badge/Docker%20Image-jsta/bathymetry-green.svg)](https://cloud.docker.com/repository/docker/jsta/bathymetry) --->

Code and data
for:

**Stachelek J.**, P. J. Hanly, P. A., Soranno. Accepted. Imperfect slope measurements drive overestimation in geometric cone model of lake and reservoir depth. Inland Waters.

### Products

Paper: [manuscript/manuscript.pdf](manuscript/manuscript.pdf)

Data: [bathymetry derived depth predictors](data/depth_predictors.csv)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
taxonomy <- read.csv("data/taxonomy.csv",
  stringsAsFactors = FALSE)

# see scripts/99_make_taxonomy.R
dt <- read.csv("data/depth_predictors.csv",
  stringsAsFactors = FALSE) %>%
  names() %>%  as.data.frame() %>%  setNames("col_name") %>%
  left_join(taxonomy) %>% data.frame() %>%
  dplyr::filter(!is.na(description)) %>%
  setNames(c("column name", "description", "variable")) %>%
  select(`column name`, variable, description)

options(knitr.kable.NA = "-")
knitr::kable(dt)
```

Data: [bathymetry derived lake depth](data/00_bathy_depth/00_bathy_depth.csv)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# see scripts/99_make_taxonomy.R
dt       <- read.csv("data/00_bathy_depth/00_bathy_depth.csv",
  stringsAsFactors = FALSE) %>%
  names() %>%  as.data.frame() %>%  setNames("col_name") %>%
  left_join(taxonomy) %>% data.frame() %>%
  dplyr::filter(!is.na(description)) %>%
  setNames(c("column name", "description", "variable")) %>%
  select(`column name`, variable, description)

knitr::kable(dt)
```

### Reproducibility

```shell
# local
make -t data/00_hypso/hypso.csv
make -t data/00_bathy_depth/bathy_pnts.rds
make -t data/00_bathy_depth/00_bathy_depth.csv
make -t data/00_bathy_depth/bathy_geometry.csv
make -t data/00_hypso/hypso_classes.csv
make -t data/00_reservoir_classification/reservoir_classes_clean.csv
make -t data/lagosus_depth_predictors.csv
make -t data/lagosne_depth_predictors.csv
make -t data/00_geometry/nearshore.csv
make all
```

<!--- ```shell --->
<!--- # TODO: docker --->
<!--- ``` --->